Wireless Automobile Detection, License Plate Processing, and Data Availability Network

### 2014-05-01 paper/paper.tex
Paper Complete? Maybe. One more good pass through should help root out any little problems

\begin{abstract}
This paper describes a parking lot monitoring system to be used at the Colorado School of Mines (CSM). The system is designed to help campus members enjoy a better parking experience by spending less time trying to find open parking lots. The project was a semester project for the computer science graduate course \emph{Wireless Sensor Networks}, taught by Dr. Tracy Camp at CSM. The work described in this paper is an extension of a project called Smartlots started by campus computing group ACMx, which uses a wireless sensor network to track the number of cars in specific parking lots. The system extends the ACMx project by adding components to the sensor network that record the license plate numbers of cars as they enter and exit lots, and by adding an administrator interface to the already existing Smartlots website. The information gathered by the sensor network is displayed on a website that lets users survey the state of the lots before arriving on campus, allowing them to make informed parking decisions. The website is also the result of collaboration between the authors of this paper and the ACMx group. The paper provides some background information, describes the parking lot monitoring system and its components, discusses the details of our implementation, and provides some initial results from testing the system and its various components.
\end{abstract}

\section{Introduction}
The parking lots at the Colorado School of Mines (CSM) can be a source of frustration for students and faculty members. Because space on campus is limited, the lots are often unable to accommodate everyone who needs to use them. Students need to choose a lot before they go to class; however, knowing which lots have free spaces usually requires driving through the lots looking for a space. The extra time spent searching one, or many, lots is frustrating, and can make students or faculty late for class. If campus members could discover which lots had available spaces prior to arrival, much time and frustration could be saved. This project involved the design and creation of a working parking lot monitoring system that will be the first step towards a better parking experience on the CSM campus. The system keeps track of both the number of vehicles within a parking lot and the license plate numbers of those vehicles. The information is made available online to system administrators and CSM campus members. There are two main parts to this project:

@@

The idea behind the first part is that a student can get on their smartphone, tablet, or laptop before heading to class, see a nicely-formatted list/map of the lots on campus, and decide where to park based on how full each lot is. Administrators can monitor the information and make sure it is correct, updating the web application as needed with new statistics, lists, and maps as more lots are added to the system. Before beginning this project, the Association for Computing Machinery (ACMx) student organization at CSM had already started working on a system for monitoring lot car counts and capacities (see section 2). We collaborated with the ACMx group, integrating our new systems and work with theirs to design and test our parking lot monitoring system.

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. Currently, our web application simply lists the license plate numbers for site administrators to view, but we envision that this monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes. The system itself consists of several subsystems, which are described in detail in section 2.

\section{Related Work}
The ACMx group at CSM has designed a system named SmartLots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station is a Raspberry Pi connected to the internet via Ethernet. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. This system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. At the time of writing, the ACMx team has created a website that will display the status of the CTLM upper and lower lots once the monitoring system has been deployed. \cite{parkingWiki} is updated periodically with new information regarding project progress. The project described in this paper augments and extends the ACMx project while collaborating with Stillwell (ACMx president and coauthor of \cite{stillwell2013}).

@@

This project is partially an extension of the ACMx SmartLots project. The system consists of several components, each of which will is either an extension of a current SmartLots system or an entirely new subsystem that has been integrated into the existing system. Figure \ref{fig:system} shows the overall architecture of our system. The system relies on three main components:
\begin{enumerate}
\item Detection Subsystem,
\item Central Base Station, and
\item Server.
\end{enumerate}

The detection subsystem consists of a small number of sensor nodes, placed strategically at parking lot entrances. Each sensor node will contain an Arduino Fio equipped with a triaxial magnetometer that will be used to detect when a car passes the sensor. The work to determine the hardware configuration and software for the Arduino Fios has already been done by the ACMx group in \cite{stillwell2013}. Each sensing node is also equipped with a Raspberry Pi single-board computer and a 5 megapixel camera, which take photographs of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer will detect the car, and the Arduino Fio will send a hardware interrupt to the Raspberry Pi, running a Python script so that the camera can take a picture of the license plate. These images are then run through an OpenCV Python program to reduce the size of the image, through downscaling, cropping, and compression algorithms.

A Raspberry Pi operates as a central base station that receives automobile detection and status data from every sensor node, via transmission from the XBee radios attached to the sensor nodes. The base station Raspberry Pi will be centrally located in a nearby building with internet connectivity. The base station utilizes its network connection to forward the data to the ACMx server.

The server is a Linode hosted machine running Ubuntu 13.10 with 2 GB RAM and 2 CPU cores. The web interface is served using Nginx. The server hosts a PHP web application with a RESTful HTTP interface that both receives the data from the base station and routes it to the correct part of the application (database, image processor, etc.). This interface is the method of communication between the server and the base station. When data is received from the base station, the application processes it, stores it in a MySQL database, and displays it appropriately on a user-facing website. An administrator website also exists that provides direct access to license plate data and car counts.

The following sections describe the architecture and function of each of the aforementioned subsystems in detail. The detection subsystem is broken up into two parts: Automobile Detection (section 3.1) and License Plate Image Acquisition (section 3.2). The glue between the sensors and the server is the Central Base Station (section 3.3). The server subsystem is also broken into two parts: Server Processing (section 3.4) and Web Application (section 3.5).

\subsection{Automobile Detection}
The automobile detection subsystem provides a means for detecting entry and departure of vehicles from a parking lot. This subsystem is designed to be placed on the side of the road next to each parking lot entrance and exit. Automobiles are detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem is based around the commercially available Arduino Fio 16-bit platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet is sent through the onboard XBee radio to the central base station (described in section 3.3). The packet  contains the direction of the passing vehicle, the processor's temperature, the battery voltage, and a portion of the magnetometer data stream. At the time of detection, a hardware interrupt is sent to the license plate recognition system to trigger it. Interrupts may also be initiated by a mechanical button that has also been included for ease of testing.

@@

The hardware used for automobile detection would be contained in the same enclosure as the license plate image acquisition hardware for simplicity. Both components share a unified power supply based on a high-capacity, 6000 mAh lithium polymer battery, which is charged by a relatively small photovoltaic panel as seen in figure \ref{fig:autodetect}. A 5V boost converter is required to interface with the Acquisition Raspberry Pi since lithium polymer batteries only provide 3.7V.

@@

\subsection{License Plate Imaging}
Once the Acquisition Raspberry Pi receives the signal from the vehicle detection Arduino Fio, it begins the process of obtaining a picture of the vehicle's license plate. This procedure is separated into two separate sections: image acquisition and image processing

\subsubsection{Image Acquisition}
While the automobile detection subsystem correctly identifies when a vehicle enters the lot, the image acquisition subsystem is responsible for determining when to capture a picture of that car in order to optimize the quality of the license plate image. The timing difference between when the Raspberry Pi receives the signal from the Arduino Fio and when the image is captured is set to 500 milliseconds so as to allow the car to completely pass the Arduino Fio and enter the capture area. Only one picture is taken of each car so that capture time and power consumption on the Raspberry Pi are limited. This frees up more resources to be allocated towards the process of image processing.

\subsubsection{Image Processing}
Once an image is captured, the Raspberry Pi creates a subprocess which processes the captured image. The image is processed using the OpenCV (Open Source Computer Vision Library) API \cite{openCV}. This API provides many different algorithms that can be used for image detection and shape recognition. It was chosen for both its ease of use and widespread cross-compatibility; it has interfaces for development in C, C++, Python, Java, and MATLAB, and supports all major operating systems (including distributions of Linux such as Raspbian).

When processing the image, two factors are taken into consideration:
\begin{enumerate}
\item File Size and
\item License Plate Quality
\end{enumerate}
The file size is reduced as much as possible in order to expedite the file transfer process from the Acquisition Raspberry Pi. The second factor, license plate quality, is considered because the image must be readable by the Optical Character Recognition (OCR) software used on the server. Several different methods of processing the images using OpenCV were attempted, including but not limited to:
\begin{enumerate}
\item Contour Mapping,
\item Image Cropping,
\item Image Resizing and Compression, and
\item Color Manipulation
\end{enumerate}

Two datasets are analyzed on how they ran through the OCR software (section 4.2). These two datasets, "cropped images" and "compressed images" had different methods of processing applied. "Cropped images" is a set of images that were naively cropped to eliminate space at the top and bottom of the picture around the car and were converted to grayscale images. "Compressed images" is a set of images that had various amounts of compression applied to them as well as different color filters.

Once the software has processed the image is sent to the ACMx server via a wi-fi connection.

@@

The central base station is the point of contact between the sensor network and the server. The data gathered by the Arduino Fio sensors are routed to the base station via XBee radios. The base station, which is itself a Raspberry Pi (see figure \ref{fig:basestation}, forwards the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station.The base station runs the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing Python scripts that take care of forwarding the data via HTTP to the server was simple. The more difficult task was writing code to allow the base station Raspberry Pi to interface with the XBee radio, and listen for packets coming from the XBee attached to the Arduino Fio. The Python code that runs on the base station Raspberry Pi is listed in Appendix 2. The base station is also equipped to receive images via the XBee radio, and forward this data to the server. However, the time constraints of the project did not allow us to finish the code on the image acquisition Raspberry Pis that sends the image data via XBees. Therefore, when testing, we simply connected the image acquisition Raspberry Pis to a WiFi internet connection, and POSTed directly to the server, using the RESTful web services already in place.

\subsection{Server Image Processing}
Once an image has been captured by an Acquisition Raspberry Pi, it is transmitted to the server via the REST API provided on the server. The code that process the image once it is received on the server serves several purposes. First, it authenticates the sender by checking for a username and password. Next, it timestamps the image and saves it to a directory on the server not directly accessible from the web. Lastly, the script calls a function to invoke two different automatic license plate recognition (ALPR) programs and stores the resulting character extractions to a MySQL database.

@@

Much of the research done on ALPR technology is less concerned with implementation and more interested in algorithmic advances. License plate recognition relies heavily on manipulation of graphics files and computer vision - each fields of study in their own right. Despite the lack of detailed literature on implemented programs, we were still able to find papers that briefly mentioned two operational systems: OpenALPR and JavaANPR. Note that ALPR and ANPR are synonymous (some packages use the term 'number' plate recognition in place of 'license' plate recognition).

@@

The two programs do share the same general algorithm for extracting license plates. First, the plate is identified using some means to isolate rectangles in the image (texture, color, boundaries, etc.). Once the bounding box containing the plate is found, segmentation begins to isolate each individual character on the license plate. Finally, these image representations of characters are mapped to their most likely plain text counterpart.

In general, ALPR systems only work for a specific geographic region they are "trained" for. For example, North American plates are a different size than European plates. Detection algorithms to identify a license plate in America would not work on an English car without modifications to the ALPR's configuration. Similarly, different states and countries use different fonts on their plates. In order to achieve maximum accuracy, you would need to collect hundreds of samples of the font and teach the program what those letters look like \cite{nikolaos2006}.

@@

One commercial provider of ALPR software is Q-Free ASA, which is based out of the Netherlands. Q-Free is a top search result when querying for ALPR software. Their ALPR product, named Intrada, offers a cloud based solution where we are able to send the image of the license plate to their servers, and get back the plain text representation of the characters. One unique feature Intrada includes that we did not find in any other package surveyed is the ability to identify the state of origin of the license plate. Intrada does this by analyzing minor differences in fonts used across state borders \cite{intrada2014}.

@@

\subsubsection{Selection of ALPR Programs}
Our literature and industry surveys left us with three viable options for license plate recognition: OpenALPR, JavaANPR, and Q-Free's Intrada ALPR. Of those three, only OpenALPR was ready to be integrated into our network. Intrada ALPR is a commercial service, so our ability to use it was dependent on Q-Free granting an educational use license to us for the semester. Lastly, JavaANPR would need to be modified to detect and recognize North American plates and fonts. JavaANPR also lacks additional information that both OpenALPR and Intrada ALPR provide such as execution time and confidence.

The goal of this project is to demonstrate a proof of concept network that is capable of monitoring parking lots and capturing license plate data. To stay aligned with this goal, and because of our unfamiliarity with graphical algorithms, we decided not to pursue using JavaANPR as a software solution for our network.

@@

Instead of choosing between two ALPR systems, we decided to integrate both platforms into our network to allow an ALPR comparison between the two tools, and offer better data integrity by checking if results from both systems (for the same photo) matched.

\subsubsection{Test Data Collection}
In order to evaluate the ALPR platforms, we needed to collect test data. Both platforms were already trained to run North American license plates out-of-the-box, so the stage was set to process images for a comparison in performance. With our final deployment in mind, we also wanted to collect images from a variety of different angles and distances, as the resulting evaluation will help determine the best camera placement in the CTLM parking lot. To accurately test the ALPR programs for our project, the test images need to be taken with the same camera that would be used during deployment.

We loaded a special testing script written in Python onto a Raspberry Pi to accomplish this task. The script was programmed to wait for a hardware interrupt before taking an image. We wired a button via a breadboard to provide the hardware interrupt. Our team brought this Pi, camera, and button out to the CTLM parking lot and positioned it near where it would be deployed.

@@

Besides images of the same license plate, we also collected several images of nearby parked cars, as well as cars that entered or exited the CTLM parking lot during that time period. And because our system is to be deployed on a college campus, we also collected images from different states, as the population of nonresidents is much greater here than in neighboring small towns.

Overall, we were able to collect 52 images of seven different plates that contained recognizable numbers. Each image from the Raspberry Pi camera was about 3 megabytes in size in its uncompressed form. The Python script also specified that the camera should save in a JPEG (.jpg) format.

@@

\subsubsection{ALPR-Project Integration}
Both OpenALPR and Intrada ALPR had to be integrated into our web application so they would run automatically without the need for human intervention. OpenALPR, because it is not a cloud service, needed to be installed locally on the ACMx server. The Intrada ALPR was much simpler because we were only interacting with the API.

\textbf{A. OpenALPR}
As stated in Section 3.4.1, OpenALPR is built on top of OpenCV, a computer vision library, and Tesseract OCR, an OCR engine. Tesseract OCR had additional dependencies that needed to be installed before it would compile.

@@

Once OpenCV and Tesseract were operational, compiling OpenALPR was a breeze. The entire process of installing all the dependencies and compiling everything (correctly) took a full working day. After installation, OpenALPR was able to start recognizing license plates immediately because it comes bundled with training data for both North American and European plates. The result of an OpenALPR output string contains
\begin{inparaenum}[\itshape a\upshape]
\item the extracted license plate characters;
\item the confidence in the characters; and
\item the execution time.
\end{inparaenum}

\textbf{B. Intrada ALPR}
Intrada ALPR uses the SOAP protocol for exchanging information \cite{soapw3c}. To create SOAP requests, we used the \verb+SoapClient+ class available in PHP to create, send, and receive the output of our request.

Intrada requires a username and password to process images. The SOAP API allows for the sending of the username and password as part of each request object for authentication. Then, Intrada runs their proprietary algorithm on its own servers before returning a result string back to our web application. The result string contains colon delimited fields that carry 
\begin{inparaenum}[\itshape a\upshape)]
\item the extracted license plate characters;
\item the confidence in the extraction;
\item the state of origin of the license plate;
\item the confidence in the state of origin;
\item coordinates of the boundary box of the detected license plate; and
\item the execution time.
\end{inparaenum}

\textbf{C. MySQL Interface}
The resulting plain text representations of the license plate are stored in MySQL database tables. There exists one table for OpenALPR recognitions, and one table for Intrada ALPR recognitions. Each table has the same structure: it stores the license plate and a reference to the 'images' table where the image filepath is stored. The Intrada ALPR table also contains a column for the recognized state. These fields are then used by the web application to be displayed to administrator users.

\subsubsection{ALPR Evaluation Environment}
We evaluated our two programs (OpenALPR and Intrada ALPR) by having each service process three different data sets: the unaltered images, cropped images, and compressed images. OpenALPR ran on the ACMx server, while Intrada ALPR ran as a cloud service, so its environment was outside of our control. Although each service offered methods to enqueue an entire directory of images and process them in bulk, we opted to evaluate them individually, as will be done in deployment.

@@

Prior to the start of our project, the ACMx group created a web application designed to display the data collected from the sensors in the parking lot monitoring system. Since the current site was essentially just a shell with no data, one main goal of this project was to improve the website. The site consists of four main parts:

\begin{enumerate}
\item Front-facing, publicly-accessible "Smartlots" website,
\item Back-end HTTP interface,
\item MySQL database, and
\item Administrator website.
\end{enumerate}

>>>>

\begin{abstract}
This paper describes a parking lot monitoring system to be used at the Colorado School of Mines (CSM). The system is designed to help campus members enjoy a better parking experience by spending less time trying to find open parking lots. The project was a semester project for the computer science graduate course \emph{Wireless Sensor Networks}, taught by Dr. Tracy Camp at CSM. The work described in this paper is an extension of a project called Smartlots started by the campus computing group ACMx, which uses a wireless sensor network to track the number of cars in specific parking lots. The system extends the ACMx project by adding components to the sensor network that record the license plate numbers of cars as they enter and exit lots, and by adding an administrator interface to the already existing Smartlots website. The information gathered by the sensor network is displayed on a website that lets users survey the state of the lots before arriving on campus, allowing them to make informed parking decisions. The website is also the result of collaboration between the authors of this paper and the ACMx group. The paper provides some background information, describes the parking lot monitoring system and its components, discusses the details of our implementation, and provides some initial results from testing the system and its various components.
\end{abstract}

\section{Introduction}
The parking lots at the Colorado School of Mines (CSM) can be a source of frustration for students and faculty members. Because space on campus is limited, the lots are often unable to accommodate everyone who needs to use them. Students need to choose a lot before they go to class; however, knowing which lots have free spaces usually requires driving through the lots looking for a space. The extra time spent searching parking lots is frustrating, and can make students or faculty late for class. If campus members could discover which lots had available spaces prior to arrival, much time and frustration could be saved. This project involved the design and creation of a parking lot monitoring system that will be the first step towards a better parking experience on the CSM campus. The system keeps track of both the number of vehicles within a parking lot and the license plate numbers of those vehicles. The information is made available online to system administrators and CSM campus members. There are two main parts to this project:

@@

The idea behind the first part is that a student can get on their smartphone, tablet, or laptop before heading to class, see a nicely-formatted list/map of the lots on campus, and decide where to park based on how full each lot is. Administrators can monitor the information and make sure it is correct, updating the web application as needed with new statistics, lists, and maps as more lots are added to the system. Before beginning this project, the Association for Computing Machinery (ACMx) student organization at CSM had already started working on a system for monitoring lot car counts and capacities (see section 2). We collaborated with the ACMx group, integrating our new systems with theirs to design and test our parking lot monitoring system.

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. Currently, our web application simply lists the license plate numbers for site administrators to view, but we envision that this monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes. The system itself consists of several subsystems, which are described in detail in section 3.

\section{Related Work}
The ACMx group at CSM has designed a system named Smartlots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station is a Raspberry Pi connected to the internet via Ethernet. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The Smartlots system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. This system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. Before we started our project, the ACMx team had created a website that would display the status of the CTLM upper and lower lots once the monitoring system had been deployed. \cite{parkingWiki} is updated periodically with new information regarding project progress. The project described in this paper augments and extends the ACMx project while collaborating with Roy Stillwell (ACMx president and coauthor of \cite{stillwell2013}).

@@

The project described in this paper is an extension of the ACMx Smartlots project. Our system consists of several components, each of which is either an extension of a previously developed Smartlots system or an entirely new subsystem that has been integrated into the existing system. Figure \ref{fig:system} shows the overall architecture of our system. The system relies on three main components:
\begin{enumerate}
\item Detection Subsystem,
\item Central Base Station, and
\item Server.
\end{enumerate}

The detection subsystem consists of a small number of sensor nodes, placed strategically at parking lot entrances. Each sensor node consists of an Arduino Fio equipped with a triaxial magnetometer that is used to detect passing vehicles. Before beginning this project, most of the work to determine the hardware configuration and software for the Arduino Fios had already been done by the ACMx group in \cite{stillwell2013}. Each sensing node is also equipped with a Raspberry Pi single-board computer and a 5 megapixel camera, which takes photographs of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer detects the vehicle, and the Arduino Fio sends a hardware interrupt to the Raspberry Pi, which kicks off a Python script so that the camera can take a picture of the vehicle's license plate. These images are then run through an OpenCV Python program to reduce the size of the image, through downscaling, cropping, and compression algorithms.

Another Raspberry Pi operates as a central base station that receives automobile detection and status data from every sensor node, via transmission from the XBee radios attached to the sensor nodes. The base station Raspberry Pi will be centrally located in a nearby building with internet connectivity. The base station utilizes its network connection to forward the data to the ACMx server.

The ACMx server is a Linode-hosted machine running Ubuntu 13.10. The web interface that displays the data we collect is served using Nginx. The Nginx server hosts a PHP web application with a RESTful HTTP API that both receives the data from the base station and routes it to the correct part of the application (database, image processor, etc.). This interface is the method of communication between the server and the base station. When data is received from the base station, the application processes it, stores it in a MySQL database, and displays it appropriately on a user-facing website. Part of the website is for administrators only, and provides direct access to license plate data and other system data that is not available to the general public.

The following sections describe the architecture and function of each of the aforementioned subsystems in detail. The detection subsystem is broken up into two parts: Automobile Detection (section 3.1) and Image Acquisition and Processing (section 3.2). The glue between the sensors and the server is the Central Base Station (section 3.3). The server subsystem is also broken into two parts: License Plate Recognition (section 3.4) and Web Application (section 3.5).

\subsection{Automobile Detection}
The automobile detection subsystem provides a means for detecting entry and exit of vehicles from a parking lot. This subsystem is designed to be placed on the side of the road next to each parking lot entrance and exit. Automobiles are detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem is based around the commercially available Arduino Fio 16-bit platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet is sent through the onboard XBee radio to the central base station (described in section 3.3). The packet  contains the direction of the passing vehicle, the processor's temperature, the battery voltage, and a portion of the magnetometer data stream. At the time of detection, a hardware interrupt is sent to the image acquisition Raspberry Pi, to notify it that a picture should be taken by its onboard camera. An easily accessible button is also wired onto the breadboard to allow a human to trigger the onboard camera for testing purposes.

@@

The hardware used for automobile detection is contained in the same enclosure as the plate image acquisition hardware for simplicity. Both components share a unified power supply based on a high-capacity, 6000 mAh lithium polymer battery, which is charged by a relatively small photovoltaic panel as seen in Figure \ref{fig:autodetect}. A 5V boost converter is required to interface with the image acquisition Raspberry Pi since the lithium polymer batteries we use only provide 3.7V.

@@

\subsection{Image Acquisition and Processing}
Once the Acquisition Raspberry Pi receives the signal from the vehicle detection Arduino Fio, it begins the process of obtaining a picture of the vehicle's license plate. This procedure is separated into two separate sections: image acquisition and image processing

\subsubsection{Image Acquisition}
While the automobile detection subsystem is responsible for correctly identifying when a vehicle enters the lot, the image acquisition subsystem is responsible for determining when to capture a picture of that car in order to optimize the quality of the license plate image. The timing difference between when the Raspberry Pi receives the signal from the Arduino Fio and when the image is captured is set to 500 milliseconds so as to allow the car to completely pass the Arduino Fio and enter the capture area. Only one picture is taken of each car so that capture time and power consumption on the Raspberry Pi are limited. This frees up more resources to be allocated towards image processing.

\subsubsection{Image Processing}
Once an image is captured, the Raspberry Pi creates a subprocess that takes care of manipulating the captured image. The image is processed using the OpenCV (Open Source Computer Vision Library) API \cite{openCV}. This API provides many different algorithms that can be used for image detection and shape recognition. It was chosen for both its ease of use and widespread cross-compatibility; it has interfaces for development in C, C++, Python, Java, and MATLAB, and supports all major operating systems (including distributions of Linux such as Raspbian).

When processing the image, two factors are taken into consideration:
\begin{enumerate}
\item File Size and
\item License Plate Quality.
\end{enumerate}
The file size is reduced as much as possible in order to expedite the file transfer process from the Acquisition Raspberry Pi. The second factor, license plate quality, is considered because the image must be readable by the Optical Character Recognition (OCR) software used on the server. Several different methods of processing the images using OpenCV were attempted, including but not limited to:
\begin{enumerate}
\item Contour Mapping,
\item Image Cropping,
\item Image Resizing and Compression, and
\item Color Manipulation.
\end{enumerate}

Two datasets were analyzed to see how well the license plate recognition software worked on various types of images (section 4.2). These two datasets, \emph{cropped images} and \emph{compressed images} had different methods of processing applied. The\emph{cropped images} were naively cropped to eliminate space at the top and bottom of the picture. These images were converted to grayscale to reduce file size. The \emph{compressed images}had various amounts of compression applied to them, as well as different color filters.

Once the OpenCV software has processed the image, it is sent from the image acquisition Raspberry Pi to the ACMx server via a Wi-Fi connection. As mentioned in section 5.4, we originally intended to send the compressed images to the base station through the same XBee radio that sends the magnetometer sensor's data, but we discovered that attempting to do this opened a can of worms that was too complicated to solve given the time constraints placed upon us.

@@

The central base station is the point of contact between the sensor network and the server. The data gathered by the Arduino Fio sensors are routed to the base station via XBee radios. The base station, which is itself a Raspberry Pi (see Figure \ref{fig:basestation}, forwards the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station, along with all the other Raspberry Pis in our system, runs the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing Python scripts that take care of forwarding the data via HTTP to the server was simple. The more difficult task was writing code to allow the base station Raspberry Pi to interface with the XBee radio, and listen for packets coming from the XBee attached to the Arduino Fio. The Python code that runs on the base station Raspberry Pi is listed in Appendix 2. The base station is also equipped to receive images via the XBee radio, and forward this data to the server. However, as previously mentioned, the time constraints of the project did not allow us to finish the sensor-side code that would need to run on the image acquisition Raspberry Pis to make image transmission via XBee possible. Therefore, when testing, we simply connected the image acquisition Raspberry Pis to a Wi-Fi internet connection, and POSTed directly to the server, using the RESTful web services already in place.

\subsection{License Plate Recognition}
Once an image has been captured by an Acquisition Raspberry Pi, it is transmitted to the server via the REST API provided on the server. The code that processes the image once it is received on the server serves several purposes. First, it authenticates the sender by checking for a username and password. Next, it timestamps the image and saves it to a directory on the server not directly accessible from the web. Lastly, the script calls a function to invoke two different automatic license plate recognition (ALPR) programs and stores the resulting character extractions to a MySQL database.

@@

Much of the research done on ALPR technology is less concerned with implementation and more interested in algorithmic advances. License plate recognition relies heavily on manipulation of graphics files and computer vision - each fields of study in their own right. Despite the lack of detailed literature on implemented programs, we were still able to find papers that briefly mentioned two operational systems: OpenALPR and JavaANPR. Note that ALPR and ANPR are synonymous (some packages use the term `number plate recognition' in place of `license plate recognition').

@@

The two programs do share the same general algorithm for extracting license plates. First, the plate is identified using some means to isolate rectangles in the image (texture, color, boundaries, etc.). Once the bounding box containing the plate is found, segmentation begins to isolate each individual character on the license plate. Finally, these image representations of characters are mapped to their most likely plain text counterparts.

In general, ALPR systems only work in the specific geographic region for which they have been `trained.' For example, North American plates are a different size than European plates. Detection algorithms to identify a license plate in America would not work on an English car without modifications to the ALPR's configuration. Similarly, different states and countries use different fonts on their plates. In order to achieve maximum accuracy, you would need to collect hundreds of samples of the fonts and teach the program what those letters look like \cite{nikolaos2006}.

@@

One commercial provider of ALPR software is Q-Free ASA, which is based in the Netherlands. Q-Free is a top search result when searching the web for ALPR software. Their ALPR product, named Intrada, offers a cloud-based solution that allows users to send the image of the license plate to their servers, and get back the plain text representation of the characters. One unique feature Intrada includes that we did not find in any other package surveyed is the ability to identify the state of origin of the license plate. Intrada does this by analyzing minor differences in fonts used across state borders \cite{intrada2014}.

@@

\subsubsection{Selection of ALPR Programs}
Our literature and industry surveys left us with three viable options for license plate recognition: OpenALPR, JavaANPR, and Q-Free's Intrada ALPR. Of those three, only OpenALPR was ready to be integrated into our network. Intrada ALPR is a commercial service, so our ability to use it was dependent on Q-Free granting an educational use license to us for the semester. Lastly, JavaANPR would need to be modified to detect and recognize North American plates and fonts. JavaANPR also lacks additional information that both OpenALPR and Intrada ALPR provide, such as execution time and confidence range.

A main goal of this project is to demonstrate a proof of concept network that is capable of monitoring parking lots and capturing license plate data. To stay aligned with this goal, and because of our unfamiliarity with graphical algorithms, we decided not to pursue using JavaANPR as a software solution for our network.

@@

Instead of choosing between two ALPR systems, we decided to integrate both platforms into our network to allow a comparison between the two tools, and offer better data integrity by checking if results from both systems (for the same photo) matched.

\subsubsection{Test Data Collection}
In order to evaluate the ALPR platforms, we needed to collect test data. Both platforms were already trained to run North American license plates out-of-the-box, so the stage was set to process images for a comparison in performance. With our final deployment in mind, we also wanted to collect images from a variety of different angles and distances, as the resulting evaluation could help determine the best camera placement for a real-world deployment. To accurately test the ALPR programs for our project, the test images needed to be taken with the same camera that would be used during a deployment.

We loaded a special testing script written in Python onto a Raspberry Pi to accomplish this task. The script was programmed to wait for a hardware interrupt before taking an image. We wired a button via a breadboard to provide the hardware interrupt. Our team brought this Pi, camera, and button out to a parking lot on campus and positioned it in a location that would be a likely candidate spot for a future deployment.

@@

Besides acquiring many images of the same license plate, we also collected several images of nearby parked cars, and of cars that entered or exited the parking lot during that time period. And because our system is designed to be deployed on a college campus, we also collected images from different states, as the population of nonresidents is much greater here than in neighboring small towns.

Overall, we were able to collect 52 images of seven different plates that contained recognizable numbers. Each image from the Raspberry Pi camera was about 3 megabytes in size in its uncompressed form. The Python script also specified that the camera should save in a JPEG format.

@@

\subsubsection{ALPR-Project Integration}
Both OpenALPR and Intrada ALPR had to be integrated into our server-side code so they would run automatically without the need for human intervention. Because OpenALPR it is not a cloud service, it needed to be installed locally on the ACMx server. The Intrada ALPR was much simpler because we were only interacting with the API.

As stated in Section 3.4.1, OpenALPR is built on top of OpenCV, a computer vision library, and Tesseract OCR, an OCR engine. Tesseract OCR had additional dependencies that needed to be installed before it would compile.

@@

Once OpenCV and Tesseract were operational, compiling OpenALPR was a breeze. The entire process of installing all the dependencies and compiling everything (correctly) took a full working day. After installation, OpenALPR was able to start recognizing license plates immediately because it comes bundled with training data for both North American and European plates. The result of an OpenALPR output string contains
\begin{inparaenum}[\itshape a\upshape)]
\item the extracted license plate characters,
\item the confidence in the characters; and
\item the execution time.
\end{inparaenum}

\textbf{B. Intrada ALPR}

Intrada ALPR uses the SOAP protocol for exchanging information \cite{soapw3c}. To create SOAP requests, we used the \verb+SoapClient+ class available in PHP to create, send, and receive the output of our request.

Intrada requires a username and password to process images. The SOAP API allows for the sending of the username and password as part of each request object for authentication. Then, Intrada runs their proprietary algorithm on its own servers before returning a result string back to our web application. The result string contains colon delimited fields that carry 
\begin{inparaenum}[\itshape a\upshape)]
\item the extracted license plate characters,
\item the confidence in the extraction,
\item the state of origin of the license plate,
\item the confidence in the state of origin,
\item coordinates of the boundary box of the detected license plate, and
\item the execution time.
\end{inparaenum}

\textbf{C. MySQL Interface}

The resulting plain text representations of the license plate are stored in MySQL database tables. There exists one table for OpenALPR recognitions, and one table for Intrada ALPR recognitions. Each table has the same structure: it stores the license plate and a reference to the 'images' table where the image filepath is stored. The Intrada ALPR table also contains a column for the recognized state. These fields are then used by the web application to be displayed to administrator users.

\subsubsection{ALPR Evaluation Environment}
We evaluated our two programs (OpenALPR and Intrada ALPR) by having each service process three different data sets: the \emph{unaltered images}, the \emph{cropped images}, and the \emph{compressed images}. OpenALPR ran on the ACMx server, while Intrada ALPR ran as a cloud service, so its environment was outside of our control. Although each service offered methods to enqueue an entire directory of images and process them in bulk, we opted to evaluate them individually, because our system would be getting images from the sensors individually in a real-world deployment.

@@

Prior to the start of our project, the ACMx group created a web application designed to display the data collected from the sensors in their Smartlots parking lot monitoring system. Since the current site was essentially just a shell with no data, one main goal of this project was to improve the website. Now that we have accomplished that goal, the new web application consists of four main parts:

\begin{enumerate}
\item Front-facing, publicly-accessible Smartlots website,
\item Back-end HTTP interface,
\item MySQL database, and
\item Administrator website.
\end{enumerate}
